The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis

### 2013-10-08 reports/Draft.Rnw
updated study overview figure
further edits... close to submission?

\doublespacing
\begin{document}
\maketitle
\begin{singlespace}
\begin{abstract}
One consequence of the expanding road network and its associated traffic is increased levels of traffic noise.  While the hedonic literature has consistently shown a negative effect of this phenomenon on the real estate market, the relationship between housing prices and automobile traffic noise has rarely been explored at landscape level, especially in the United States. Here, we show this relationship by modeling the propagation of traffic noise over the landscape and by analyzing the extent of its impact on single-family home transactions in the urban area surrounding St.\ Paul, Minnesota using spatially explicit local regression techniques. Our estimate of the negative relationship between traffic noise levels and house prices is smaller in magnitude than most of the relevant published literature. We find no evidence that the effect of traffic noise varies significantly by geography, nor by traffic noise levels.  We find significant differences in the impact of traffic noise before and after the economic recession of 2008-09. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September 2008, and an average of 0.37 percent after September 2008.
\end{abstract}
\end{singlespace}

\section{Introduction and Past Research}\label{sec:lit}
Prolonged exposure to traffic noise affects people in a number of ways, ranging from simple annoyance \citep{Miedema2001, Ouis2001, Ohrstrom2007, DeKluizenaar2013, Weinhold2013}, to sleep disturbance \citepalias{Netherlands2004}, to increasing risk for stroke \citep{Sorensen2011}, hypertension \citep{Jarup2008, Bodin2009}, myocardial infarction \citep{Babisch2005}, and overall quality of life \citep{Shepherd2013}. The noise level at which such effects are observed does not have to be high.  It has been shown that people exposed to traffic noise with a 24-hour average of 55 decibels (dBA) are found to be at a higher risk for hypertension \citep{Barregard2009, Bodin2009}, and those exposed to 60 dBA or greater are found to be at a higher risk for stroke \citep{Sorensen2011}.

@@

Policies at the local, state, and federal level can influence the level and effect of traffic noise. For instance, the US Federal Highway Admistration reports that 47 states have constructed noise barriers to reduce noise propagation, spending over a half a billion dollars from 2008-2010 \citepalias{USFHA2012}. The efficiency of such projects requires an understanding of the economic cost of traffic noise.

@@

Reviews of the research by \citet{Bateman2001} found that the negative effect of traffic noise on house prices ranged from 0.08 to 2.22 percent per decibel with a mean around 0.55 percent per decibel, while \citet{Navrud2002} found an average decrease of 0.64 percent per decibel. Much of the research in this literature has tended to focus on the European experience with noise, especially the United Kingdom \citep{Day2007, Blanco2011}, Sweden \citep{Wilhelmsson2000, Andersson2010}, Switzerland \citep{Baranzini2010}, Spain \citep{MarmolejoDuarteCarlos;GonzalezTamez2009}, Netherlands \citep{Theebe2004a}, and Germany \citep{Brandt2011}.

Instead of focusing on the impact of automobile traffic, work in the United States has tended to concentrate on noise from airports \citep{Espey2000, McMillen2004, Cohen2008a} or indirect effects of traffic noise, such as proximity to highways \citep{Matthews2007, Chernobai2009, Li2012}, or traffic counts \citep{HughesJr.1992, Larsen2012}. One exception is \citet{Cheng2008}, which, using micro data in Louisville, TN during 2007, finds that an additional decibel of traffic noise reduced house prices by 0.34 percent. However, this research is based on less than a thousand transactions in just one city.

\citet{Delucchi1998} used an estimate of 0.85 percent reduction in house prices per additional decibel in their cost-benefit analysis of traffic noise. While their preferred estimate of the total damage cost to the United States in 1991 is \$3 billion, they report a possible range of less than \$100 million to over \$40 billion due in part to uncertainty surrounding the degree of noise propagation over space in the urban built environment and the effect of noise on house prices. Indeed, one of the main conclusions of their work is that more research should be done to better understand the impact of noise on house prices and that better noise data are needed for such work to occur.

One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling traffic noise propagation over the landscape. To properly analyze the spatial association between real estate prices and exposure to traffic noise at a landscape level, it is necessary to create a noise surface map at sufficiently detailed spatial resolution to account for the complex and heterogeneous interaction between the noise source and the resistance of the landscape to noise propagation.  Implementing such a model can be very difficult.  For one thing, the data needed for the model is very extensive and may not even be readily available (e.g., building footprint and height data).  Furthermore, it is computationally very intensive. Fortunately, recent developments in Geographic Information Systems and distributive computing have reduced these difficulties, making it much easier to create a noise surface map at landscape level with high spatial resolution.

In this paper, we contribute to the noise hedonic literature by examining the relationship between single family home prices from 2005 to 2010 and variation in road traffic noise exposure in the St.\ Paul, Minnesota urban area across 44 different cities. First, we utilize new estimates of traffic noise that have never before been used in the hedonic literature. The noise propagation model estimates traffic noise levels at a high spatial resolution (10m x 10m) while also incorporating the effects of nearby buildings and landcover.

We also construct a flexible hedonic model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work \citep[see][]{MarmolejoDuarteCarlos;GonzalezTamez2009, Carruthers2010, Sunding2010, Nappi-Choulet2011} and have been shown to be more accurate than other spatial econometric models under common circumstances \citep{McMillen2012}. \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} used a LWR model to estimate the impact of traffic noise on housing prices in Barcelona, Spain and found evidence of a non-stationary relationship over the study area. We find that while the impact of other important control variables in the hedonic function vary over space, the impact of traffic noise does not.

Lastly, we present the first estimates of the implicit price of traffic noise after the housing market collapse and Great Recession of 2008-09. Contrary to work such as \citet{Cho2011b}, which found that environmental amenities mattered less after the housing crash, our estimated effect of traffic noise mattered more after the recession (0.37 percent reduction in house prices per additional decibel) compared to beforehand (0.19 percent per decibel).

\section{Data}
The 2010 US Census lists the population of the Twin Cities Metropolitan Region (Minneapolis and St. Paul and their surrounding areas)  as almost 3 million residents spread over seven counties. This study examines single family residential home transactions in the Census-defined urban areas of three of the seven counties: Dakota, Ramsey and Washington County (see Figure \ref{fig:overview}). We obtained data from approximately forty thousand sales transactions between 2005 and 2010 (n=42,095) from the 2010 MetroGIS Regional Parcel Dataset published by the Metropolitan Council. Figure \ref{fig:overview} shows an overview of the study area as well as the spatial distribution of the house sales prices. In addition to the geographic location and date of the house sale, we collected or calculated structural and locational variables commonly used in the hedonic literature. Table \ref{tab:sumStats} provides a brief description of the variables in our data and some basic summary statistics. Table \ref{tab:cor} displays a simple correlation matrix of the quantitative variables.

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = 1.2\textwidth]{../graphs/StudyArea_SaleValueDistribution}}
\caption{This figure shows the spatial extent of our study area as well as the significant spatial variation in single family house sales prices.}\label{fig:overview}
\end{figure}

@@

\subsection{Structural Attributes}
According to a review by \cite{Wilhelmsson2000}, the most common structural attributes included in real-estate hedonic pricing studies are living area, number of bathrooms, age, garage and lot size.  Although the 2010 MetroGIS Regional Parcel Dataset includes structural data on living area, age, garage presence, lot size, owner-occupancy and house architectural style for every transaction, we have variables for the number of bedrooms, bathrooms, and size of the garage for house sales in one county. We feel confident in our results even without these independent variables for most of our study data because sensitivity analysis conducted in areas with the additional structural variables revealed very similar estimates when the variables were included and excluded. These results are in section \ref{Discussion} of the the paper. Additionally, other hedonic work has been published using a similar set of explanatory variables in this area, see for instance, \citet{Sander2009b}.

\subsection{Noise Data}
Should we cite previous work that uses these data?

To determine the relationship between real estate price and road traffic noise levels, we first created a traffic noise exposure surface for the entire Twin Cities Metro Region (TCMR) by calculating the propagation of traffic noise over the landscape using the FHWA (Federal Highway Authority) 1978 standard \citep{Barry1978}. Following \citet{Barry1978}, the traffic noise level at any given location on the landscape is given by: 
\begin{eqnarray}\label{eq:noise}
L_{EQ}(i) &=& \bar{L}_0(i) + 0.115 \sigma _i^2 + 10 log \frac{N_i \pi D_0}{T*S_i} + 10 log \left[ \frac{D_0}{D}\right]^{1 + \alpha}  \nonumber \\
&& + 10 log \left( \frac{\psi _{\alpha (\phi _1, \phi _2)}}{\pi}\right) + \Delta _{gradient} + \Delta _{shielding}
\end{eqnarray}
where $L_{EQ}(i)$ is A-weighted hourly energy equivalent noise level in dBA, which is calculated for each class $i$ of vehicle (automobile and trucks); $\bar{L}_0(i)$ is the mean Sound Pressure Level (SPL) at the reference distance for class $i$; $\sigma _i$ is the standard deviation of the SPL for each class of vehicle; $N_i$ is the number of vehicles of the $i^{th}$ class passing during the relevant hour; $D_0$ is the reference distance (usually 15 m); $D$ is the perpendicular distance from the road center line to the receiver; $\alpha$ is a site parameter (soft and hard surface), $0 < \alpha < 1$; $S_i$ is the mean speed of the $i^{th}$ class; $T$ is the duration, usually 1 hour; $\phi _1$ and $\phi _2$ are the angles from the perpendicular of the limits of the observer's view of a section of the road. They are used to account for only the energy coming from a portion of the roadway; $\Delta _{gradient}$ is an adjustment for road surface gradient; $\Delta _{shielding}$ is a shielding adjustment (land cover, buildings, noise barriers).

To calculate the traffic noise map for the region using the above model, several data sources were used. A 2007 road centerline and the associated traffic characteristics (volume and proportion of trucks and vehicles) for the region were obtained from the Minnesota Department of Transportation (MNDoT). We converted the posted speed of each road into a GIS layer. For all roads without recorded traffic volume (i.e., residential), we assigned 100 vehicles/day as a minimum estimate. This estimate was based on factors identified in the literature as important for estimating traffic volume, such as census data on commuting population, tract area, the size of a standard street block, and road type \citep{Cheng1992, Fricker1986}.\footnote{In order to assess the sensitivity of our results to our estimate of traffic volume for local roads, we also ran the model using 200 and 300 vehicles/day for one of the counties that make up the TCMR. With the exception of a few areas in the rural part of the county, there was less than 0.05 dBA difference at 30 m from the road centerline.}

@@

The locations chosen for the traffic noise levels were the 818,500 buildings in the TCMR. Thus, we used the perimeter and height of these buildings to account for the shielding effect of buildings. To account for foliage shielding (and to make the computation manageable), we assumed that only forest patches that are at least 10 m high and have an area greater than 10,000 m$^2$ will have a significant effect on noise attenuation. Accordingly, we extracted forest polygons with these characteristics using a combination of year 2005 land use data and 2001 National Land Cover Data (NLCD). To account for noise barrier shielding, we used a noise barrier layer obtained from MNDoT, which contains the locations, width, and height of noise barriers in the region.

We used the SoundPlan$^{TM}$ noise modeling software to implement the model given by Equation \ref{eq:noise}. Predicted traffic noise output is calculated at a grid resolution of 10 m$^2$.\footnote{We conducted a preliminary validation of the model output by comparing it with observed noise levels at 134 locations that were sampled along major highways. For each location, two readings were taken very close to the highway, one between 9am and 10am and another between 1pm and 2pm. The noise level for each location is then determined by averaging the two observed values, and this average is then compared with the predicted noise level. The relationship between the mean observed and predicted noise levels is linear and moderately strong with a correlation of 0.76. It is not surprising that the relationship is not stronger given the size and complexity of the model. In particular, the traffic volume data that we used for modeling traffic noise propagation is a 24-hour average, which requires more time-interval sampling per location than just the two used in the validation. Still, this preliminary comparison serves as a good indicator of how well this model captures traffic noise levels.} Once the noise surface for the entire region was developed, we extracted the noise surface for the present study area by overlaying the housing parcel boundaries and calculating the maximum noise level within the parcel.

\subsection{Other Locational Attributes}
A common real estate addage states that the three most important real estate attributes are: location, location and location. Knowing where each house is located allows us to also construct a vector of other attributes associated with the sales transaction. For instance, using GIS software we are able to calculate the Euclidean distance to numerous points of interest within the dataset, such as the nearest central business district, shopping centers, parks, and lakes. Additionally, a variable denoting the median household income for the surrounding census tract is created through the use of TIGER shapefiles from the 2010 Census Bureau and data from the 2010 American Community Survey (ACS). Lastly, we associate each transaction with its elementary school and include the average 3rd grade Minnesota Comprehensive Assessment (MCA) score for the local elementary school during the year of purchase.\footnote{Test scores were obtained from the Minnesota Department of Education website. The school district and elementary school attendance boundary spatial information was taken from the Minnesota Geospatial Information Office Clearinghouse Data Catalog.}

@@

\begin{table}
\begin{center}
\caption{Mean Variable Values by Year}\label{tab:SumStatsTime}
\small
%\includegraphics[width = \textwidth]{../graphs/DescriptiveStatsByYear}
% Table generated by Excel2LaTeX from sheet 'Descriptive StatisticsYear'
    \begin{tabular}{lrrrrrr|r}
    Variable & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2005-10\\ \hline
    Sales Value (\$1000s) & 279   & 282   & 275   & 263   & 231   & 233   & 266 \\
    House Size (ft$^2$) & 1737  & 1736  & 1743  & 1813  & 1727  & 1775  & 1750 \\
    Lot Size (acres) & 0.25  & 0.25  & 0.25  & 0.26  & 0.25  & 0.26  & 0.25 \\
    Year Built  & 1967  & 1966  & 1966  & 1971  & 1968  & 1968  & 1967 \\
    Traffic Noise (dB) & 58.5  & 58.6  & 58.6  & 52.5  & 52.8  & 52.8  & 56.4 \\
    Median Income (\$1000s) & 72.0  & 72.1  & 72.3  & 74.8  & 72.6  & 73.2  & 72.6 \\
    Elem. Test Scores & 364.8 & 364.8 & 365.2 & 363.5 & 362.3 & 366.1 & 364.5 \\
    Dist. to CBD (km) & 14.7  & 14.6  & 14.8  & 15.0  & 14.3  & 14.5  & 14.7 \\
    Dist. to Park (km) & 2.6   & 2.6   & 2.6   & 2.8   & 2.6   & 2.6   & 2.6 \\
    Dist. to Lake (km) & 0.9   & 0.9   & 0.9   & 0.9   & 0.9   & 0.9   & 0.9 \\
    Dist. to Shop (km) & 1.9   & 1.9   & 1.8   & 2.0   & 1.9   & 1.9   & 1.9 \\ \hline
    Number of Observations &  10,991  & 8,885  &  6,549  &  5,503  &  5,961  & 4,206  & 42,095  \\
    \end{tabular}%
\end{center}
\end{table}

\section{Basic Econometric Model}\label{basicModel}
Consistent with past research, this study implements a semi-logarithmic hedonic pricing model. Our aim is to estimate the marginal willingness to pay for different attributes, in particular changes in the traffic noise associated with the house. Equation~\eqref{eq:model} expresses the basic hedonic model.
\begin{equation}\label{eq:model}	
ln \textrm{ Sale Price}_i = \beta _0 + \beta _1 \textrm{Noise}_i+ \beta _2 S_i+ \beta _3 N_i + \beta _4 T_i + \textrm{error}_i
\end{equation}
Where Noise$_i$ is the noise level for house $i$, $S_i$ is a vector of the house's structural attributes, $N_i$ is a vector of the neighborhood attributes, and $T_i$ is a vector of time fixed effects. We can interpret the regression model coefficients as the price semi-elasticities of the underlying attributes. For instance, we can interpret the coefficient on noise as the percentage increase in price for a one decibel increase in the traffic noise associated with the transaction in our dataset.

@@

Table \ref{tab:mainresults} presents the results of our preferred LWR model as well as a basic global OLS model for comparison. The first two columns display the estimates and standard errors from a basic semi-log OLS regression with traffic noise, structural characteristics, location-based amenities, and both city and year of sale dummy variables. The third column displays the mean estimated coefficients from the LWR model. In this model coefficients are estimated at each observation within the dataset using the nearest 200 house sales within the past 12 months. The fourth column displays the standard deviation of the estimated coefficients to give a sense of how the coefficients vary when allowed to do so through the use of locally weighted regression techniques. In addition to the significant reduction in the GCV score (0.041 vs. 0.027) obtained with the use of LWR, we also see a substantal reduction in the degree of spatial autocorrlation in the model residuals. We calculate the Moran’s I statistic to be 0.199 for the global model while our prefered LWR specification reduces the Moran's I statistic to 0.012, a nearly twenty-fold reduction.

Column (5) in Table \ref{tab:mainresults} shows the percentage of the local regressions for which the coefficient's p-value is less than 0.1. All regressions contained statistically significant estimates of the impact of increased living space and a majority contained significant estimates for lot size and year of construction. Roughly half of the regressions estimated an impact of traffic noise statistically different from zero, while approximately one-third of the local regressions estimated statistically significant estimates of the locational variables (school test scores, census tract median income, distance to local amenities).

>>>>

\doublespacing
\begin{document}
\maketitle
\begin{singlespace}
\begin{abstract}
One consequence of the expanding road network and its associated traffic is increased levels of traffic noise.  While the hedonic literature has consistently shown a negative effect of this phenomenon on the real estate market, research in the United States has often relied on crude measures of traffic noise. Here, we reduce the measurement error of traffic noise exposure through a detailed model of noise propagation over the landscape. Additionally, we estimate the impact on single-family home transactions throughout the St.\ Paul, Minnesota, urban area using spatially explicit local regression techniques to allow for spatial non-stationarity in the hedonic function. Contrary to previous work, we find no evidence that the effect of traffic noise varies significantly across the several hundred square miles of our study region, nor by traffic noise levels. We do find significant differences in the impact of traffic noise before and after the economic recession of 2008-09. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September 2008 and an average of 0.37 percent after September 2008.
\end{abstract}
\end{singlespace}

\newpage
\section{Motivation and Past Research}\label{sec:lit}
Prolonged exposure to traffic noise affects people in a number of ways, ranging from simple annoyance \citep{Miedema2001, Ouis2001, Ohrstrom2007, DeKluizenaar2013, Weinhold2013}, to sleep disturbance \citepalias{Netherlands2004}, to increasing risk for stroke \citep{Sorensen2011}, hypertension \citep{Jarup2008, Bodin2009}, myocardial infarction \citep{Babisch2005}, and overall quality of life \citep{Shepherd2013}. The noise level at which such effects are observed does not have to be high.  It has been shown that people exposed to traffic noise with a 24-hour average of 55 decibels (dBA) are found to be at a higher risk for hypertension \citep{Barregard2009, Bodin2009}, and those exposed to 60 dBA or greater are found to be at a higher risk for stroke \citep{Sorensen2011}.

@@

@@

Reviews of the research by \citet{Bateman2001} found that the negative effect of traffic noise on house prices ranged from 0.08 to 2.22 percent per decibel with a mean around 0.55 percent per decibel, while \citet{Navrud2002} found an average decrease of 0.64 percent per decibel. Much of the recent research in this literature has tended to focus on the European experience with noise, often due to the availability of high quality traffic noise data made available from the government, especially the United Kingdom \citep{Day2007, Blanco2011}, Sweden \citep{Wilhelmsson2000, Andersson2010}, Switzerland \citep{Baranzini2010}, Spain \citep{MarmolejoDuarteCarlos;GonzalezTamez2009}, Netherlands \citep{Theebe2004a}, and Germany \citep{Brandt2011}.

Instead of focusing on the impact of automobile traffic, hedonic work in the United States has tended to concentrate on noise from airports \citep{Espey2000, McMillen2004, Cohen2008a} or indirect effects of traffic noise, such as proximity to highways \citep{Matthews2007, Chernobai2009, Li2012}, or traffic counts \citep{HughesJr.1992, Larsen2012}. One exception is \citet{Cheng2008}, which finds that an additional decibel of traffic noise reduced house prices in Louisville, Tennessee, by 0.34 percent during the year 2007. However, the generalizability of this research is limited given that it is based on less than a thousand house transactions in just one city.

A major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling traffic noise propagation over the landscape. To properly analyze the spatial association between real estate prices and exposure to traffic noise, it is necessary to create a noise surface map at sufficiently detailed spatial resolution to account for the complex and heterogeneous interaction between the noise source and the resistance of the landscape to noise propagation.  Implementing such a model can be very difficult.  The data needed for the model is very extensive and may not even be readily available (e.g., building footprint and height data).  Furthermore, it is computationally very intensive. Fortunately, recent developments in Geographic Information Systems and distributive computing have reduced these difficulties, making it much easier to create a noise surface map at landscape level with high spatial resolution.

In this paper, we estimate the relationship between house prices and traffic noise across 44 cities and several hundred square miles in the St.\ Paul, Minnesota, urban region by taking advantage of uniquely detailed set of noise estimates that have never before been used in the hedonic literature. The data represent perhaps the most accurate regional noise estimates ever used in the hedonic literature in the United States to date due to the high spatial resolution of the estimates (10 m x 10 m) and the incorporation of the effects of nearby buildings and landcover on noise propagation. More detail about the noise model inputs and procedures are presented in the next section of the paper.

%Precise estimates of the impact of traffic noise are important for efficient implementation of noise mitigation projects. The US Federal Highway Admistration reports that 47 states have constructed noise barriers to reduce noise propagation, spending over a half a billion dollars from 2008-2010 \citepalias{USFHA2012}. \citet{Delucchi1998} used an estimate of 0.85 percent reduction in house prices per additional decibel in their cost-benefit analysis of traffic noise. While their preferred estimate of the total damage cost to the United States in 1991 is \$3 billion, they report a possible range of less than \$100 million to over \$40 billion due in part to uncertainty surrounding the degree of noise propagation over space in the urban built environment and the effect of noise on house prices. Indeed, one of the main conclusions of their work is that more research should be done to better understand the impact of noise on house prices and that better noise data are needed for such work to occur.

Work from Europe suggests that the relationship between traffic noise and house prices varies across housing submarkets \citep{Day2007, MarmolejoDuarteCarlos;GonzalezTamez2009}, by level of traffic noise \citep{Theebe2004a, Brandt2011}, and over time \cite{Wilhelmsson2000}. As such, we construct a flexible hedonic model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work \citep[see][]{MarmolejoDuarteCarlos;GonzalezTamez2009, Carruthers2010, Sunding2010, Nappi-Choulet2011} and have been shown to be more accurate than other common spatial econometric models under common real world circumtances \citep{McMillen2012}. Thus, in addition to using new data that reduces the measurement error of the impact of traffic noise, the spatial and temporal variation in our data and hedonic modeling allow us to present the first estimates in the United States of how the impact of traffic noise varies over space, time, and by noise level.

We find that while the impact of other important control variables in the hedonic function vary significantly over space, the impact of traffic noise does not. Such results are contradictory to recent work by \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} which found evidence of a spatially non-stationary relationship between traffic noise and housing prices using a LWR model in Barcelona, Spain. Our paper also presents the first estimates of the implicit price of traffic noise after the housing market collapse and economic recession of 2008-09. Contrary to work such as \citet{Cho2011b}, which found that environmental amenities mattered \emph{less} after the housing crash, our estimated effect of traffic noise mattered \emph{almost twice as much more} after the recession (0.19 percent reduction in house prices per additional decibel beforehand compared to 0.37 percent per decibel afterwards). We conclude with a discussion of how future research can expand and improve upon this work.

\section{Study Area and Data}
The 2010 US Census lists the population of the Twin Cities Metropolitan Region (Minneapolis and St. Paul and their surrounding areas) as almost 3 million residents spread over seven counties. This study examines single family residential home transactions in the Census-defined urban areas of three of the seven counties: Dakota, Ramsey and Washington County (see Figure \ref{fig:overview}). We obtained sales data from approximately forty thousand transactions between 2005 and 2010 (n=42,095) from the 2010 MetroGIS Regional Parcel Dataset published by the Metropolitan Council. Figure \ref{fig:overview} shows an overview of the study area as well as the spatial distribution of the house sales prices.

In addition to the geographic location and date of the house sale, we collected or calculated structural and locational variables commonly used in the hedonic literature, such as the size of the house, lot size, and architectural style of the house, as well as distances to amenities like the Central Business District, parks, and lakes. Table \ref{tab:sumStats} provides a brief description of the variables in our data and some basic summary statistics. Table \ref{tab:cor} displays a simple correlation matrix of the quantitative variables.

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = 1.2\textwidth]{../graphs/StudyAreaSaleValueDist}}
\caption{This figure shows the spatial extent of our study area as well as the significant spatial variation in single family house sales prices.}\label{fig:overview}
\end{figure}

@@

\subsection{Noise Data}
To determine the relationship between real estate price and road traffic noise levels, we first created a traffic noise exposure surface by calculating the propagation of traffic noise over the landscape using the FHWA (Federal Highway Authority) 1978 standard \citep{Barry1978}. Following \citet{Barry1978}, the traffic noise level at any given location on the landscape is given by: 
\begin{eqnarray}\label{eq:noise}
L_{EQ}(i) &=& \bar{L}_0(i) + 0.115 \sigma _i^2 + 10 log \frac{N_i \pi D_0}{T*S_i} + 10 log \left[ \frac{D_0}{D}\right]^{1 + \alpha}  \nonumber \\
&& + 10 log \left( \frac{\psi _{\alpha (\phi _1, \phi _2)}}{\pi}\right) + \Delta _{gradient} + \Delta _{shielding}
\end{eqnarray}
where $L_{EQ}(i)$ is A-weighted hourly energy equivalent noise level in decibels (dBA), which is calculated for each class $i$ of vehicle (automobile and trucks); $\bar{L}_0(i)$ is the mean Sound Pressure Level (SPL) at the reference distance for class $i$; $\sigma _i$ is the standard deviation of the SPL for each class of vehicle; $N_i$ is the number of vehicles of the $i^{th}$ class passing during the relevant hour; $D_0$ is the reference distance (usually 15 m); $D$ is the perpendicular distance from the road center line to the receiver; $\alpha$ is a site parameter (soft and hard surface), $0 < \alpha < 1$; $S_i$ is the mean speed of the $i^{th}$ class; $T$ is the duration, usually 1 hour; $\phi _1$ and $\phi _2$ are the angles from the perpendicular of the limits of the observer's view of a section of the road. They are used to account for only the energy coming from a portion of the roadway; $\Delta _{gradient}$ is an adjustment for road surface gradient; $\Delta _{shielding}$ is a shielding adjustment (land cover, buildings, noise barriers).

To calculate the traffic noise map for the region using the above model, several data sources were used. A 2007 road centerline and the associated traffic characteristics (volume and proportion of trucks and vehicles) for the region were obtained from the Minnesota Department of Transportation (MNDoT). We converted the posted speed of each road into a GIS layer. For all roads without recorded traffic volume (i.e., residential), we assigned 100 vehicles/day as a minimum estimate. This estimate was based on factors identified in the literature as important for estimating traffic volume, such as census data on commuting population, tract area, the size of a standard street block, and road type \citep{Cheng1992, Fricker1986}.\footnote{In order to assess the sensitivity of our results to our estimate of traffic volume for local roads, we also ran the model using 200 and 300 vehicles/day for one county. With the exception of a few areas in the rural part of the county, there was less than 0.05 dBA difference at 30 m from the road centerline.}

@@

The locations chosen for the traffic noise levels were the 818,500 buildings in the entire urban area. Thus, we used the perimeter and height of these buildings to account for the shielding effect of buildings. To account for foliage shielding (and to make the computation manageable), we assumed that only forest patches that are at least 10 m high and have an area greater than 10,000 m$^2$ will have a significant effect on noise attenuation. Accordingly, we extracted forest polygons with these characteristics using a combination of year 2005 land use data and 2001 National Land Cover Data (NLCD). To account for noise barrier shielding, we used a noise barrier layer obtained from MNDoT, which contains the locations, width, and height of noise barriers in the region.

We used the SoundPlan$^{TM}$ noise modeling software to implement the model given by Equation~\eqref{eq:noise}. Predicted traffic noise output is calculated at a grid resolution of 10 m$^2$.\footnote{We conducted a preliminary validation of the model output by comparing it with observed noise levels at 134 locations that were sampled along major highways. For each location, two readings were taken very close to the highway, one between 9am and 10am and another between 1pm and 2pm. The noise level for each location is then determined by averaging the two observed values, and this average is then compared with the predicted noise level. The relationship between the mean observed and predicted noise levels is linear and moderately strong with a correlation of 0.76. It is not surprising that the relationship is not stronger given the size and complexity of the model. In particular, the traffic volume data that we used for modeling traffic noise propagation is a 24-hour average, which requires more time-interval sampling per location than just the two used in the validation. Still, this preliminary comparison serves as a good indicator of how well this model captures traffic noise levels.} Once the noise surface for the entire region was developed, we extracted the noise surface for the present study area by overlaying the housing parcel boundaries and calculating the maximum noise level within the parcel.

\subsection{Structural Attributes}
According to a review by \cite{Wilhelmsson2000}, the most common structural attributes included in real-estate hedonic pricing studies are living area, number of bathrooms, age, garage and lot size.  Although the 2010 MetroGIS Regional Parcel Dataset includes structural data on living area, age, garage presence, lot size, owner-occupancy and house architectural style for every transaction, we only have variables for the number of bedrooms, bathrooms, and size of the garage for house sales in one county. We feel confident in our results even without these independent variables for most of our study data because sensitivity analysis conducted in areas with the additional structural variables revealed very similar estimates when the variables were included and excluded. Section \ref{Discussion} provides a more detailed description of these results. %Additionally, other hedonic work has been published using a similar set of explanatory variables in this area, see for instance, \citet{Sander2009b}.

\subsection{Other Locational Attributes}
A common real estate addage states that the three most important real estate attributes are: location, location and location. Knowing where each house is located allows us to also construct a vector of other attributes associated with the sales transaction. For instance, using GIS software we are able to calculate the Euclidean distance to numerous points of interest within the dataset, such as the nearest central business district, shopping centers, parks, and lakes. Additionally, a variable denoting the median household income for the surrounding census tract is created through the use of TIGER shapefiles from the 2010 Census Bureau and data from the 2010 American Community Survey. Lastly, we associate each transaction with its elementary school and include the average 3rd grade Minnesota Comprehensive Assessment (MCA) score for the local elementary school during the year of purchase.\footnote{Test scores were obtained from the Minnesota Department of Education website. The school district and elementary school attendance boundary spatial information was taken from the Minnesota Geospatial Information Office Clearinghouse Data Catalog.}

@@

\begin{table}
\begin{center}
\caption{Mean Variable Values by Year}\label{tab:SumStatsTime}
\small
%\includegraphics[width = \textwidth]{../graphs/DescriptiveStatsByYear}
% Table generated by Excel2LaTeX from sheet 'Descriptive StatisticsYear'
    \begin{tabular}{lrrrrrr|r}
    Variable & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2005-10\\ \hline
    Sales Value (\$1000s) & 279   & 282   & 275   & 263   & 231   & 233   & 266 \\
    Traffic Noise (dB) & 58.5  & 58.6  & 58.6  & 52.5  & 52.8  & 52.8  & 56.4 \\
    House Size (ft$^2$) & 1737  & 1736  & 1743  & 1813  & 1727  & 1775  & 1750 \\
    Lot Size (acres) & 0.25  & 0.25  & 0.25  & 0.26  & 0.25  & 0.26  & 0.25 \\
    Year Built  & 1967  & 1966  & 1966  & 1971  & 1968  & 1968  & 1967 \\
    Median Income (\$1000s) & 72.0  & 72.1  & 72.3  & 74.8  & 72.6  & 73.2  & 72.6 \\
    Elem. Test Scores & 364.8 & 364.8 & 365.2 & 363.5 & 362.3 & 366.1 & 364.5 \\
    Dist. to CBD (km) & 14.7  & 14.6  & 14.8  & 15.0  & 14.3  & 14.5  & 14.7 \\
    Dist. to Park (km) & 2.6   & 2.6   & 2.6   & 2.8   & 2.6   & 2.6   & 2.6 \\
    Dist. to Lake (km) & 0.9   & 0.9   & 0.9   & 0.9   & 0.9   & 0.9   & 0.9 \\
    Dist. to Shop (km) & 1.9   & 1.9   & 1.8   & 2.0   & 1.9   & 1.9   & 1.9 \\ \hline
    Number of Observations &  10,991  & 8,885  &  6,549  & 5,503  &  5,961  & 4,206  & 42,095  \\
    \end{tabular}%
\end{center}
\end{table}

\section{Basic Econometric Model}\label{basicModel}
Our aim is to estimate the marginal willingness to pay for different attributes, in particular changes in the traffic noise associated with the house. Consistent with past research, this study implements a semi-logarithmic hedonic pricing model.  Before implementing the Locally Weighted Regression model, we present the results of a simpler econometric model given by equation~\eqref{eq:model}.
\begin{equation}\label{eq:model}	
ln \textrm{ Sale Price}_i = \beta _0 + \beta _1 \textrm{Noise}_i+ \beta _2 S_i+ \beta _3 N_i + \beta _4 T_i + \textrm{error}_i
\end{equation}
Noise$_i$ is the noise level for house $i$, $S_i$ is a vector of the house's structural attributes, $N_i$ is a vector of the neighborhood attributes, and $T_i$ is a vector of time fixed effects. Due to the semi-logarithmic functional form of the model, we can interpret the regression model coefficients as the price semi-elasticities of the underlying attributes. For instance, we can interpret the coefficient on noise as the percentage increase in price for a one decibel increase in the traffic noise associated with the transaction in our dataset.

@@

The first two columns of Table \ref{tab:mainresults} display the estimates and standard errors from a basic semi-log OLS regression with traffic noise, structural characteristics, location-based amenities, and both city and year of sale dummy variables. The third and fourth columns summarize the results of the LWR model when coefficients are estimated at each observation within the dataset using the nearest 200 house sales within the past 12 months. The third column presents the mean of these estimates while the fourth column displays the standard deviation of the estimates to give a measure of variation in the estimates across the dataset. Most of the mean LWR coefficients, including traffic noise, are similar to the coefficients obtained with the global OLS model. However, the large standard deviations of the LWR coefficients suggest that there is significant variation in the estimated coefficients with the more flexible model. We now try to determine the source of this variation.

Column (5) in Table \ref{tab:mainresults} shows the percentage of the local regressions for which the coefficient's p-value is less than 0.1. All regressions contained statistically significant estimates of the impact of increased living space and a majority contained significant estimates for lot size and year of construction. Roughly half of the regressions estimated an impact of traffic noise statistically different from zero, while approximately one-third of the local regressions estimated statistically significant estimates of the locational variables (school test scores, census tract median income, distance to local amenities). The lack of statistical significance for many of the local regression coefficients could be due to multiple factors. First, because we are estimating local regressions with only the nearest 200 data points, rather than tens of thousands of sales as in the global model, we might be estimating the coefficients with less precision and are unable to differentiate the coefficients from zero. Alternatively, the coefficients themselves might vary, sometimes being close to zero and other times farther away. This variation might be due to random chance (after all, we have estimated tens of thousands of regressions), or it could be the result of true spatial non-stationarity on the part of the coefficients.
